Mapping mutational effects along the evolutionary landscape of HIV envelope

### 2018-03-22 paper/manuscript.tex
added results on cell-line validation

We conducted the low MOI viral passage illustrated in \FIG{dms_schematic}A in SupT1.CCR5 cells~\citep[obtained from Dr. James Hoxie;][]{boyd2015mutations}.
During this passage, cells were maintained in R10 media, which has the same composition as the D10 described above, except RPMI-1640 (GE Healthcare Life Sciences, SH30255.01) is used in the place of DMEM.
In addition, the media contained 10 $\mu$g/mL DEAE-dextran to enhance viral infection.
We infected cells with 4 million (for replicate 1) or 5 million (for replicates 2 and 3) TZM-bl infectious units of mutant virus at an MOI of 0.01, with cells at a starting concentration of 1 million cells/mL in vented tissue-culture flasks (Fisher Scientific, 14-826-80).
At day 1 post-infection, we pelleted cells, aspirated the supernatant, and resuspended cell pellets in the same volume of fresh media still including the DEAE-dextran.
At 2 days post-infection, we doubled the volume of each culture with fresh media still including DEAE-dextran.
At 4 days post-infection, we pelleted cells, passed the viral supernatant through a 0.2 $\mu$m filter, concentrated the virus $\sim$30 fold using ultracentrifugation as described in \citet{dingens2017comprehensive}, and froze aliquots at -80$^{\circ}$C.
In parallel, for each replicate, we also passaged $2\times 10^5$ (for replicate 1) or $5 \times 10^5$ (for replicates 2 and 3) TZM-bl infectious units of wildtype virus using the same procedure.
To obtain final titers for our concentrated virus, we thawed one of the aliquots stored at -80$^{\circ}$C and titered using the TZM-bl assay in the presence of 10 $\mu$g/mL DEAE-dextran.

>>>>

We conducted the low MOI viral passage illustrated in \FIG{dms_schematic}A in SupT1.CCR5 cells~\citep[obtained from Dr. James Hoxie;][]{boyd2015mutations}.
The SupT1.CCR5 cells tested negative for mycoplasma. 
The SupT1.CCR5 cell line was previously created by engineering the parental SupT1 cell line to express CCR5~\citep{boyd2015mutations}. 
We used antibody staining followed by flow cytometry to validate that our stock of SupT1.CCR5 cells expressed CCR5, CXCR4, and CD4. 
There is no validated STR profile for SupT1.CCR5 cells. 
However, we performed STR profiling on our stock of cells and compared the results to the ATCC SupT1 (ATCC \#CRL-1942) reference profile. 
We found that 11 of 14 alleles plus both amelogenin alleles matched the reference, with no additional mismatched alleles ins the SupT1.CCR5 profile. 
Given the known instability of lymphoma cell lines~\citep{inoue2000frequent}, this level of identity suggests that the SupT1.CCR5 cells are indeed related to the parental SupT1 cells~\citep{capes2013match}.

During this passage, cells were maintained in R10 media, which has the same composition as the D10 described above, except RPMI-1640 (GE Healthcare Life Sciences, SH30255.01) is used in the place of DMEM.
In addition, the media contained 10 $\mu$g/mL DEAE-dextran to enhance viral infection.
We infected cells with 4 million (for replicate 1) or 5 million (for replicates 2 and 3) TZM-bl infectious units of mutant virus at an MOI of 0.01, with cells at a starting concentration of 1 million cells/mL in vented tissue-culture flasks (Fisher Scientific, 14-826-80).
At day 1 post-infection, we pelleted cells, aspirated the supernatant, and resuspended cell pellets in the same volume of fresh media still including the DEAE-dextran.
At 2 days post-infection, we doubled the volume of each culture with fresh media still including DEAE-dextran.
At 4 days post-infection, we pelleted cells, passed the viral supernatant through a 0.2 $\mu$m filter, concentrated the virus $\sim$30 fold using ultracentrifugation as described in \citet{dingens2017comprehensive}, and froze aliquots at -80$^{\circ}$C.
In parallel, for each replicate, we also passaged $2\times 10^5$ (for replicate 1) or $5 \times 10^5$ (for replicates 2 and 3) TZM-bl infectious units of wildtype virus using the same procedure.
To obtain final titers for our concentrated virus, we thawed one of the aliquots stored at -80$^{\circ}$C and titered using the TZM-bl assay in the presence of 10 $\mu$g/mL DEAE-dextran.

### 2018-03-22 paper/manuscript_for_final_submission/manuscript_nofigs_embeddedtabsrefs.tex
reference later GitHub commit
added results on cell-line validation

@@

\bibitem[{Haddox et~al.(2018)Haddox, Hugh K and Dingens, Adam S and Hilton,
  Sarah K and Overbaugh, Julie and Bloom, Jesse D}]{haddox2018github}
\textbf{\color{eLifeMediumGrey} Haddox HK}, Dingens AS, Hilton SK, Overbaugh J,
  Bloom JD.
\newblock Computer code for ``Mapping mutational effects along the evolutionary
  landscape of HIV envelope''.
\newblock Github.  2018; commit ec358fd.
\newblock
  \urlprefix\url{https://github.com/jbloomlab/EnvMutationalShiftsPaper}.

@@

>>>>

\bibitem[{Capes-Davis et~al.(2013)Capes-Davis, Amanda and Reid, Yvonne A and
  Kline, Margaret C and Storts, Douglas R and Strauss, Ethan and Dirks, Wilhelm
  G and Drexler, Hans G and MacLeod, Roderick AF and Sykes, Gregory and Kohara,
  Arihiro and others}]{capes2013match}
\textbf{\color{eLifeMediumGrey} Capes-Davis A}, Reid YA, Kline MC, Storts DR,
  Strauss E, Dirks WG, Drexler HG, MacLeod RA, Sykes G, Kohara A, et~al.
\newblock Match criteria for human cell line authentication: where do we draw
  the line?
\newblock International Journal of Cancer.  2013; 132(11):2510--2519.

@@

\bibitem[{Haddox et~al.(2018)Haddox, Hugh K and Dingens, Adam S and Hilton,
  Sarah K and Overbaugh, Julie and Bloom, Jesse D}]{haddox2018github}
\textbf{\color{eLifeMediumGrey} Haddox HK}, Dingens AS, Hilton SK, Overbaugh J,
  Bloom JD.
\newblock Computer code for ``Mapping mutational effects along the evolutionary
  landscape of HIV envelope''.
\newblock Github.  2018; commit 2355e76.
\newblock
  \urlprefix\url{https://github.com/jbloomlab/EnvMutationalShiftsPaper}.

@@

\bibitem[{Inoue et~al.(2000)Inoue, Kaoru and Kohno, Takashi and Takakura,
  Satoshi and Hayashi, Yasuhide and Mizoguchi, Hideaki and Yokota,
  Jun}]{inoue2000frequent}
\textbf{\color{eLifeMediumGrey} Inoue K}, Kohno T, Takakura S, Hayashi Y,
  Mizoguchi H, Yokota J.
\newblock Frequent microsatellite instability and BAX mutations in T cell acute
  lymphoblastic leukemia cell lines.
\newblock Leukemia Research.  2000; 24(3):255--262.
